\nomenclature{$\pfami$}{The parametric family of the component lifetimes.}
\nomenclature{$\pfam$}{The parametric family of the system.}
\nomenclature{$\infomatrx(\tparam\sysparamvec \given w)$}{The \emph{expected} information matrix for a random masked system failure times conditioned on $w$ candidates.}
\nomenclature{$\infomatrxn(\tparam\sysparamvec \given w)$}{The \emph{expected} information matrix for a random sample of $n$ masked system failure times conditioned on $w$ candidates.}
\nomenclature{$\infomatrx(\tparam\sysparamvec)$}{The \emph{expected} information matrix for a random masked system failure times.}
\nomenclature{$\infomatrxn(\tparam\sysparamvec)$}{The \emph{expected} information matrix for a random sample of $n$ masked system failure times.}
\nomenclature{$\tparam\sysparamvec$}{The true parameter index (vector) of the parametric family of the system.}
\nomenclature{$\tparam\sysparam$}{The true parameter index (matrix) of the parametric family of the system.}
\nomenclature{$\mleu$}{The maximum likelihood estimator conditioned on a sample of $n$ masked system failure times.}
\nomenclature{$\mle$}{The maximum likelihood estimator conditioned on a sample of $n$ masked system failure times in which only candidate sets of cardinality $w$ are observed
.}
\nomenclature{$\mfs$}{A random sample of $n$ masked system failure times.}
\nomenclature{$\mfso$}{A realization of a sample of $n$ masked system failure times.}
\nomenclature{$\mfsw$}{A random sample of $n$ masked system failure times conditioned on $w$ candidates.}
\nomenclature{$\RV{W}$}{The discrete random variable denoting the cardinality of the candidate set.}
\nomenclature{$\cv_j$}{The continuous random lifetime of the \jth component.}
\nomenclature{$\sv$}{The continuous random lifetime of the series system.}
\nomenclature{$\RV{K}$}{The discrete random variable denoting the cause of system failure.}
\nomenclature{$\rvcand$}{The discrete random variable (set) denoting the components of a candidate set.}
\nomenclature{$\score(\sysparamvec \given w\,,\,\cdot\;)$}{The score function, the gradient of the log-likelihood.}
\nomenclature{$\ell(\sysparamvec \given w\,,\,\cdot\;)$}{The log-likelihood function conditioned on $\RV{W} = w$.}
\nomenclature{$\likelihood(\sysparamvec \given w\,,\,\cdot\;)$}{The likelihood function.}
\nomenclature{$\cv_j$}{The continuous random lifetime of the \jth component.}
\nomenclature{$\pdf_j$}{The probability density function of $\cv_j$.}
\nomenclature{$\spdf$}{The probability density function of $\sv$.}
\nomenclature{$\jointpdf$}{The joint conditioned probability density function of $\rvcand$, $\sv$ given some realization of $\RV{W}$.}
\nomenclature{$\cov(\tparam\sysparamvec \given w)$}{The \emph{expected} variance-covariance of the asymptotic sampling distribution of $\mle$.}
\nomenclature{$\cov(\tparam\sysparamvec)$}{The \emph{expected} variance-covariance of the asymptotic sampling distribution of $\mleu$.}
\nomenclature{$\pmf_{\RV{W}}$}{The marginal probability mass function of $\rv{W}$.}
\nomenclature{$\pmf_{\RV{K} \given \sv, \rv{W}}$}{The conditional probability mass function of $\rv{W}$ given some realization of $\sv$ and $\rv{W}$.}
\nomenclature{$\sysparam$}{A parameter index (matrix) of the parametric family of the system.}
\nomenclature{$\sysparamvec$}{A parameter index (vector) of the parametric family of the system.}
\nomenclature{$\sysparam_j$}{A parameter index (vector) of the parametric family of the \jth component.}
\nomenclature{$\cands$}{The sample space of $\rvcand$, the set of all candidate 
sets.}
\nomenclature{$\candsw{w}$}{The sample space of $\rvcand$ given $\RV{W} = w$, 
the set of all candidate sets of cardinality $w$.}
\nomenclature{$\indicator{\mathcal{X}}$}{The indicator function 
$\indicator{\mathcal{X}} \colon \mathcal{X} \mapsto \{0,1\}$ outputs $1$ if the 
input is a member of $\mathcal{X}$ and otherwise outputs $0$.}
\nomenclature{$\sampd$}{The sampling distribution of $\mle$.}
\nomenclature{$\sampdu$}{The sampling distribution of $\mleu$.}
