Next, we consider what occurs when we increase the masking probability $p$ to $0.333$, which represents a fairly significant masking of the component cause of failure.
```{r samp-size-n-vs-stats-p333-scale, fig.cap=c("Sample Size vs Bootstrapped Scale CI Statistics (p = 0.333)", "Sample Size (n) vs Coverage Probability, IQR of Bootstrapped CIs and 95% Confidence Band of Scale MLEs for p = 0.333"), fig.align="center", echo = F}
knitr::include_graphics("image/plot-n-vs-stats-p333-scale.pdf")
```

Figure \ref{fig:samp-size-n-vs-stats-p333-scale} shows the distribution of the MLEs for the shape parameters of the first three components and the bootstrapped CIs for different sample sizes with a component cause of failrue masking probaility of $p = 0.333$ (each
non-failed component is in the candidate set with a $33.3\%$ probabiltiy).

Here are several key observations, similar to the previous scenario:

- *Coverage Probability (CP)*: The CP is well-calibrated, obtaining a value near the nominal 95% level across different sample sizes.
                               Again, this suggests that the bootstrapped CIs will contain the true value of the shape parameter with the specified confidence level. In general,
                               the CIs are neither too wide nor too narrow, but the CP for $\lambda_1$ at the sample size of 50 is relatively lower at 0.88. This might indicate a bit
                               of uncertainty at smaller sample sizes for this parameter.
                               
    According to Section \ref{sec:reliability}, the MTTF of component 1 is the largest, and so it is
    the least likely to fail. This might explain the lower CP for $\lambda_1$ at smaller sample sizes, as the data has not shown enough failures from
    component 1 to accurately estimate its scale parameter. If we were to increase its MTTF to a much larger value, it would almost never
    fail, and the uncertainty about the scale parameter would further increase.

- *Dispersion of MLEs*: As with the previous plots, the shaded regions, which represents the 95% probability range of the MLEs, narrows with incresing sample sizes.
                        This indicates an increase in precision when more data is available. If this did not occur, then it would indicate that the MLEs are not consistent
                        or are likelihood model is not appropriate for the data.

- *IQR of Bootstrapped CIs*: The IQR (vertical blue bars) reduces with an increase in sample size. This suggests that the bootstrapped CIs are getting more consistent and focused around a narrower range with larger samples while maintaining a good coverage probability. As we get more data, the bootstrapped CIs are more likely to be closer to each other and the true value of the scale parameter.

- *Mean of MLEs*: The red dashed line indicating the mean of MLEs is less stable across different sample sizes, but it is still close to the true value.
                  We will investigate the effect of the masking probability in more detail in Section \ref{sec:p-vs-mttf}.

### Shape Parameters

```{r samp-size-n-vs-stats-p215-shape, fig.cap=c("Sample Size vs Bootstrapped Shape CI Statistics (p = 0.215)", "Sample Size (n) vs Coverage Probability, IQR of Bootstrapped CIs and 95% Confidence Band of Shape MLEs for p = 0.333"), fig.align="center", echo = F}
knitr::include_graphics("image/plot-n-vs-stats-p215-shape.pdf")
```

Figure \ref{fig:samp-size-n-vs-stats-p215-shape} shows the distribution of the MLEs for the shape parameters of the first three components and the bootstrapped CIs for different sample sizes with a component cause of failrue masking probaility of $p = 0.333$ (each
non-failed component is in the candidate set with a $33.3\%$ probabiltiy).

Here are several key observations, similar to the previous scenario:

- *Coverage Probability (CP)*: The CP is well-calibrated, obtaining a value near the nominal 95% level across different sample sizes. Again, this suggests that the bootstrapped CIs will contain the true value of the shape parameter with the specified confidence level. The CIs are neither too wide nor too narrow.

- *Dispersion of MLEs*: The shaded regions representing the 95% probability range of the MLEs get narrower as the sample size increases. This is an indicator of the increased precision in the estimates as more data is available. 

- *IQR of Bootstrapped CIs*: The IQR (vertical blue bars) reduces with an increase in sample size. This suggests that the bootstrapped CIs are getting more consistent and focused around a narrower range with larger samples while maintaining a good coverage probability. As we get more data, the bootstrapped CIs are more likely to be closer to each other and the true value of the scale parameter.

- *Mean of MLEs*: The red dashed line indicating the mean of MLEs is less stable across different sample sizes, but it is still close to the true value.
                  We will investigate the effect of the masking probability in more detail in Section \ref{sec:p-vs-mttf}.


```{r samp-size-n-vs-stats-p333-shape, fig.cap=c("Sample Size vs Bootstrapped Shape CI Statistics (p = 0.333)", "Sample Size (n) vs Coverage Probability, IQR of Bootstrapped CIs and 95% Confidence Band of Shape MLEs for p = 0.333"), fig.align="center", echo = F}
knitr::include_graphics("image/plot-n-vs-stats-p333-shape.pdf")
```

