### Scenario #4: Assessing the Impact of Changing the Shape Parameter of a Component {#sec:shape-vs-mttf}

When the shape parameters in a series system are approxiamtely equal, as is the case in our
base series system defined in Table \ref{tab:series-sys}, then a larger scale parameter yields
a larger MTTF and smaller probability of component failure. However, when the shape parameters
are not equal, then the relationship between the scale parameter and the MTTF is more complex.

There are two effects taking place in the presence of masking and right-censoring. In the case
of right-censoring, we want to increase the reliability $R_{\T_i}(\tau)$ to make it more likely
that the system is still functioning at time $\tau$. This can be accomplished by increasing the








The shape parameter of the Weibull distribution has a non-linear relationship to its MTTF.
The non-linearity complicates any analysis, and so instead we use a more relevant definition of
reliability for a series system, the probability of a component being the cause of failure.
The probability that the $j$\textsuperscript{th} component is the cause of failure is given by
$$
\Pr\{K_i = j\} = \int_{0}^\infty f_{\T_i, K_i}(t, j ; \v\theta) dt.
$$

We consider $m=3$ Weibull components in a series configuration. We will vary the shape
parameter of component 1, $k_1$, and observe the bias in the scale and shape parameters
for components 1 and 2.
The shape and scale parameters are given by
\begin{equation}
\begin{aligned}
    0.1 \leq k_1 &< 0.55    &\quad \lambda_1 = 1\\
             k_2 &= 0.5     &\quad \lambda_2 = 1\\
             k_3 &= 0.5     &\quad \lambda_3 = 1.
\end{aligned}
\end{equation}

```{r table-3, table.attr = "style='width:50%;'", echo = F, results = 'asis', fig.cap = "Component Reliability", fig.align = "center"}
mttf <- gamma(1 + 1 / c(0.2, 0.5, 0.5)) * c(1, 1, 1)
components <- data.frame(
    "MTTF" = mttf,
    "Failure Probability" = wei_series_cause(1L:3L, scales = c(1, 1, 1), shapes = c(0.2, 0.5, 0.5)),
    row.names = paste("Component", 1:3)
)
names(components) <- gsub("\\.", " ", names(components))
knitr::kable(components, caption = "Mean Time To Failure (MTTF) and Probability of Component Failure of Weibull Components in Series Configuration")
```

As shown by Table 3, the MTTF of component 1 can be significantly larger than the MTTF of the other components,
but the probability of component 1 being the cause of failure is significantly larger. In fact, as the MTTF
of component 1 increases as we decrease its shape parameter, the probability of component 1 being the cause of
failure increases. This is because the shape parameter of component 1 is less than 1, and so the hazard function
decreases with respect to time.
This means that component 1 is more likely to fail early in its lifetime (high infant mortality) and less likely
to fail later in its lifetime (low aging). This means that component 1 is more likely to be the cause of failure
than the other components, even though it has a larger MTTF.

```{r n100-p0-q1, fig.cap=c("Probability of Component 1 Failure vs Bias (p = 0, q = 1)"), fig.align="center", echo = F}
knitr::include_graphics("image/loess_smoothed_n_100_p_0_q_1_probs_0.0_-_0.55.pdf")
```

Figure \ref{fig:n100-p0-q1} depict the bias of the scale and shape parameters of components 1
and 2 in relation to the failure probability of component 1 in the series system when we
have a sample with no right censoring and no masking of the component cause of failure.
This represents an ideal scenario.
The green dashed line represents the true value of the parameters, and the red line shows the
bias.

The bias (red line) shows a non-linear behavior as the probability
of component 1 failure changes. For both the scale and shape parameters, the bias fluctuates
and does not remain constant.

It is difficult to understand the pattern of the bias for the shape and scale parameters as the
probability of component 1 failure changes. We expected the MLE of the shape parameter
to become more positively biased as the probability of component 1 failure increases in order
to nudge its MTTF down. Similarly, we expected the MLE of the scale parameter to become more
negatively biased as the probability of component 1 failure increases in order to nudge its
MTTF down. However, we observe the exact opposite. The bias of the shape and scale parameters
for component 2 are more in line with what we expected for component 1. For the same reason,
for component 2, we expected to see the bias of the shape parameter become more negative
and bias of the scale parameter to become more positive as the probability of component 1 failure
increases in order to nudge the MTTF of component 2 upwards, but we see the exact opposite.


To understand these graphs, we must better understand the relationship between the probability
of component 1 failure and the shape parameter of the components.

In the following analysis, we fix the shape parameters of components 1 and 2 to $0.5$ ($k_2 = k_3 = 0.5$)
and vary the shape parameter of component 1 ($k_1$) to observe how the probability of component failure
changes in relation to the shape parameter of component 1 ($k_1$). We also show the MTTF of the components
and the system.

Let $P_j$ denote the probability that the $j$\textsuperscript{th} component is the cause of the system
failure ($\Pr\{K_i = j\}$), where we use the notation $K_i$ to denote the component cause of failure
for the $i$\textsuperscript{th} system. Then,
$$
    P_j = \int_{0}^\infty f_{\T_i, K_i}(t, j ; \v\theta) dt =
        E_{\v\theta}\biggl\{ \frac{h_j(\T_i;\v{\theta_j})}{h_{\T_i}(\T_i;\v\theta)} \biggr\}.
$$
Since this is a complex integral, we use numerical integration to compute the probability of component
failure for different values of $k_1$.

In Table \ref{tab:table-vary-k1}, we show these results for different values of $k_1$:

```{r table-vary-k1, echo = FALSE}
shapes <- seq(0.1, 1, by = 0.1)
P1 <- rep(NA, length(shapes))
P2 <- rep(NA, length(shapes))
P3 <- rep(NA, length(shapes))
mttf1 <- rep(NA, length(shapes))
mttf23 <- rep(NA, length(shapes))
mttfsys <- rep(NA, length(shapes))

for (i in 1:length(shapes)) {
    P1[i] <- wei_series_cause(1, scales = rep(1, 3), shapes = c(shapes[i], 0.5, 0.5))
    P2[i] <- wei_series_cause(2, scales = rep(1, 3), shapes = c(shapes[i], 0.5, 0.5))
    P3[i] <- wei_series_cause(2, scales = rep(1, 3), shapes = c(shapes[i], 0.5, 0.5))
    mttf1[i] <- wei_mttf(scale = 1, shape = shapes[i])
    mttfsys[i] <- wei_series_mttf(scales = rep(1, 3), shapes = c(shapes[i], 0.5, 0.5))
}

# make a table of these values
df <- data.frame(k1 = shapes, P1 = P1, P2 = P2, P3 = P3, MTTF1 = mttf1, MTTF2 = 2, MTTF3 = 2, MTTF = mttfsys)
colnames(df) <- c("$k_1$", "$P_1$", "$P_2$", "$P_3$", "MTTF$_1$", "MTTF$_2$", "MTTF$_3$", "System MTTF")
knitr::kable(df,
    format = "latex",
    caption = "Probability of Component Failures and Mean Time To Failures As We Vary the Shape Parameter of Component 1",
    digits = 2, escape = FALSE
) %>%
    kable_styling(latex_options = "striped", full_width = FALSE)
```


Here are some key observations:

- As $k_1$ increases, $\Pr\{K_i = 3\}$ decreases. This may seem unexpected, as it *decreases* the MTTF of component $j$.
  However, this is because the shape parameter of component 1 is less than 1, and so the hazard function decreases
  with respect to time. This means that component 1 is more likely to fail early in its lifetime (high infant mortality)
  and less likely to fail later in its lifetime (low aging). This means that component 1 is more likely to be the cause
  of failure than the other components, even though it has a larger MTTF.

- In order to make it more likely to see component 1 as the cause of system failure, therefore we must decrease
  $k_1$. In Figure \ref{fig:n100-p0-q1}, the MLE nudges $k_1$ down. This is what we see in the the figure, as the bias
  decreases (the MLE for $k_1$ is nudged down).

- Conversely, if you want to nudge the probability of component 2 and 3 not being a cause of component failure,
  you should increase $k_2$ and $k_3$. This is what we see in Figure \ref{fig:n100-p0-q1} for the bias of $k_2$,
  where the MLE is nudging its bias upwards. Alternatively, you can keep it the same while only decreasing the
  shape parameter for component 1. This may indicate that the MLE is more sensitive to the shape parameter of
  component 1 than the shape parameter of component 2.

To analyze the behavior of the scale parameter bias, we show the effect of varying the scale parameter of component
1 on the probability of component failure and the MTTF of the components and the system in
Table \ref{tab:table-vary-lam1}.

```{r table-vary-lam1, echo = FALSE, eval=F}
lambda <- seq(1, 4, by = 1)
P1 <- rep(NA, length(lambda))
P2 <- rep(NA, length(lambda))
P3 <- rep(NA, length(lambda))
P4 <- rep(NA, length(lambda))
P5 <- rep(NA, length(lambda))
mttf1 <- rep(NA, length(lambda))
mttfsys <- rep(NA, length(lambda))

for (i in 1:length(lambda)) {
    scales_i <- c(lambda[i], 1, 1)
    P1[i] <- wei_series_cause(1, scales = scales_i, shapes = shapes)
    P2[i] <- wei_series_cause(2, scales = c(lambda[i], 1, 1), shapes = shapes)
    P3[i] <- wei_series_cause(3, scales = c(lambda[i], 1, 1), shapes = shapes)
    P4[i] <- wei_series_cause(4, scales = c(lambda[i], 1, 1), shapes = shapes)
    P5[i] <- wei_series_cause(5, scales = c(lambda[i], 1, 1), shapes = shapes)
    mttf1[i] <- wei_mttf(scale = lambda[i], shape = shapes[1])
    mttfsys[i] <- wei_series_mttf(scales = c(lambda[i], 1, 1), shapes = shapes)
}

mttf23 <- wei_mttf(scale = 1, shape = shapes[2])

# make a table of these values
df <- data.frame(
    lambda1 = lambda,
    P1 = P1, P2 = P2, P3 = P3, MTTF1 = mttf1, MTTF2 = mttf23, MTTF3 = mttf23, MTTF = mttfsys
)
colnames(df) <- c("$\\lambda_1$", "$P_1$", "$P_2$", "$P_3$", "MTTF$_1$", "MTTF$_2$", "MTTF$_3$", "System MTTF")
knitr::kable(df,
    format = "latex",
    caption = "Probability of Component Failures and Mean Time To Failures As We Vary the Scale Parameter of Component 1",
    digits = 2, escape = FALSE
) %>%
    kable_styling(latex_options = "striped", full_width = FALSE)
```

Here are some key observations about Table \ref{tab:table-vary-lam1}:

- The relationship is more linear and intuitive when we vary the scale parameter of component 1.
  As the scale parameter of component 1 increases, the probability of component 1 being the cause
  of failure decreases, and the probability of component 2 and 3 being the cause of failure
  increases. This is also in alignment with the reasoning that, in this case, increasing the MTTF
  of a component decreases its probability of being the cause of failure. However, when we vary
  the shape parameter of component 1, the relationship is more complex, as previously discussed.

- To increase the probability of component 1 being the cause of failure, we must decrease the
  scale parameter of component 1. This is not, however, what we see in Figure \ref{fig:n100-p0-q1}.
  The MLE nudges the scale parameter of component 1 upwards (increasing bias). It could be due to
  a conflict between nudging the scale parameter and nudging the shape parameter. It may be
  a complex linear relationship the merits further investigation. This is what we see in the the figure, as the bias
  decreases (the MLE for the scale parameter of component 1 is nudged down).

