---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T,message=F,warning=F)
library(printr)
options(digits=4)
library(series.system.estimation.masked.data)
library(tidyverse)
library(md.tools)
library(algebraic.mle)
```

R package `series.system.estimation.masked.data` {-}
====================================================

The source code for the R package `series.system.estimation.masked.data` can be
found at https://github.com/queelius/series_system_estimation_masked.data.
There is built-in help on how to use the R package.
For instance, to get help on `md_bernoulli_candidate_C1_C2_C2`, use the following R command:
```{r}
?md_bernoulli_candidate_C1_C2_C3
```

In what follows, we reproduce some of the R code found in the R package.
For a deeper look, please look at the official documentation.

`md_loglike_weibull_series_C1_C2_C3` generates the log-likelihood function for
masked data with component lifetimes that are Weibull distributed and candidate
sets that satisfy the conditions.

Here is the documentation for `md_loglike_weibull_series_C1_C2_C3`:
```{r,echo=F}
?md_loglike_weibull_series_C1_C2_C3
```

Here is a reproduction of its R source code:
```{r,echo=F}
md_loglike_weibull_series_C1_C2_C3
```

R package `algebraic.mle` {-}
=============================
The source code for the R package `algebraic.mle` can be
found at https://github.com/queelius/algebraic.mle.
There is built-in help on how to use the R package.
For instance, to get help on `confint.mle`, use the following R command
(we have also produced its output):
```{r}
library(algebraic.mle)
?confint.mle
```

We reproduce the R code for `mle_iterative`, the general iterative MLE method
described in Section \ref{sec:iterative}.

```{r}
?mle_iterative
```

```{r}
mle_iterative
```


R package `md.tools` {-}
========================
The source code for the R package `md.tools` can be
found at https://github.com/queelius/md.tools.
There is built-in help on how to use the R package.
For instance, to get help on `md_decode_matrix`, use the following R command
(we have also produced its output):
```{r}
library(md.tools)
?md_decode_matrix
```
