---
title: "Components with weibull distributed lifetimes"
output:
    bookdown::pdf_document2:
        toc: false
        number_sections: true
        keep_tex: true
        latex_engine: xelatex
        citation_package: natbib
date: "2023-04-29"
---


\newcommand{\T}{T}
\renewcommand{\v}[1]{\boldsymbol{#1}}
\numberwithin{equation}{section}


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Components with Weibull distributed lifetimes {#sec:weibull}
=============================================
Consider a series system in which the components have Weibull distributed
lifetimes.
The $j$\textsuperscript{th} component of the $i$\textsuperscript{th} has a
lifetime distribution given by
$$
    T_{i j} \sim \operatorname{WEI}(\v{\theta_j})
$$
where $\v{\theta_j} = (k_j, \lambda_j)$ for $j=1,\ldots,m$.
Thus, $\v\theta = (\v{\theta_1},\ldots,\v{\theta_m})' = \bigl(k_1,\lambda_1,\ldots,k_m,\lambda_m\bigr)$.
The random variable $T_{i j}$ has a reliability function, pdf, and hazard function
given respectively by
\begin{align}
    R_j(t;\lambda_j,k_j)
        &= \exp\biggl\{-\biggl(\frac{t}{\lambda_j}\biggr)^{k_j}\biggr\},\\
    f_j(t;\lambda_j,k_j)
        &= \frac{k_j}{\lambda_j}\biggl(\frac{t}{\lambda_j}\biggr)^{k_j-1}
        \exp\biggl\{-\left(\frac{t}{\lambda_j}\right)^{k_j} \biggr\},\\
    h_j(t;\lambda_j,k_j) \label{eq:weibull_haz}
        &= \frac{k_j}{\lambda_j}\biggl(\frac{t}{\lambda_j}\biggr)^{k_j-1}
\end{align}
where $t > 0$, $\lambda_j > 0$ is the scale parameter and $k_j > 0$ is the
shape parameter.

```{r echo=F, message=F, warning=F}
library(ggplot2)
library(dplyr)
shape1 <- 3
scale1 <- 30
shape2 <- 1
scale2 <- 50
shape3 <- 2.5
scale3 <- 15
shape4 <- 5
scale4 <- 20
shape5 <- 0.25
scale5 <- 50

ts <- seq(0,100,by=.1)
h.wei1 <- Vectorize(function(t) (shape1/scale1)*(t/scale1)^(shape1-1))
h.wei2 <- Vectorize(function(t) (shape2/scale2)*(t/scale2)^(shape2-1))
h.wei3 <- Vectorize(function(t) (shape3/scale3)*(t/scale3)^(shape3-1))
h.wei4 <- Vectorize(function(t) (shape4/scale4)*(t/scale4)^(shape4-1))
h.wei5 <- Vectorize(function(t) (shape5/scale5)*(t/scale5)^(shape5-1))

df.haz <- data.frame(
    t=ts,
    y=c(h.wei1(ts),h.wei2(ts),h.wei3(ts),h.wei4(ts),h.wei5(ts)),
    fun=rep(c("WEI(3,30)","WEI(1,50)","WEI(2.5,15)","WEI(5,20)","WEI(0.25,50)"),
        each=length(ts)))
```



```{r exp_weib_haz, fig.align="center", fig.cap="Hazard plots"}
ggplot(df.haz, aes(x=t, y=y, color=fun)) +
    geom_line() +
    labs(y="hazard") +
    ylim(0,100)
```

The shape parameter $k$ may be understood in the following way:

- If $k < 1$, then the hazard function decreases with respect to time, which may
occur if defective items fail early and are weeded out.
- If $k > 1$, then the hazard function is increases with respect
to time, which may occur as a result of an aging process.
- If $k = 1$, then the failure rate is constant, which means it is exponentially
distributed.

The lifetime of the series system composed of $m$ Weibull components
has a reliability function given by
\begin{equation}
\label{eq:sys_weibull_reliability_function}
R(t;\v\theta) = \exp\biggl\{-\sum_{j=1}^{m}\biggl(\frac{t}{\lambda_j}\biggr)^{k_j}\biggr\}.
\end{equation}
\begin{proof}
By Theorem \ref{thm:sys_reliability_function},
$$
R(t;\v\theta) = \prod_{j=1}^{m} R_j(t;\lambda_j,k_j).
$$
Plugging in the Weibull component reliability functions obtains the result
\begin{align*}
R(t;\v\theta)
    &= \prod_{j=1}^{m} \exp\biggl\{-\biggl(\frac{t}{\lambda_j}\biggr)^{k_j}\biggr\}\\
    &= \exp\biggl\{-\sum_{j=1}^{m}\biggl(\frac{t}{\lambda_j}\biggr)^{k_j}\biggr\}.
\end{align*}
\end{proof}

```{r exp_weib_pdf, cache=T, fig.cap="Density plots", echo=F, fig.width=5,fig.asp=0.8, message=F, warning=F, eval=F}
R.wei1 <- Vectorize(function(t) exp(-(t/scale1)^shape1))
f.wei1 <- Vectorize(function(t) shape1/scale1*(t/scale1)^(shape1-1)*
    exp(-(t/scale1)^shape1))
R.wei2 <- Vectorize(function(t) exp(-(t/scale2)^shape2))
f.wei2 <- Vectorize(function(t) shape2/scale2*(t/scale2)^(shape2-1)*
    exp(-(t/scale2)^shape2))
R.wei3 <- Vectorize(function(t) exp(-(t/scale3)^shape3))
f.wei3 <- Vectorize(function(t) shape3/scale3*(t/scale3)^(shape3-1)*
    exp(-(t/scale3)^shape3))
R.wei4 <- Vectorize(function(t) exp(-(t/scale4)^shape4))
f.wei4 <- Vectorize(function(t) shape4/scale4*(t/scale4)^(shape4-1)*
    exp(-(t/scale4)^shape4))
R.wei5 <- Vectorize(function(t) exp(-(t/scale5)^shape5))
f.wei5 <- Vectorize(function(t) shape5/scale5*(t/scale5)^(shape5-1)*
    exp(-(t/scale5)^shape5))
df <- data.frame(
    t=ts,
    y=c(f.exp(ts),f.wei(ts),f.wei2(ts)),
    fun=rep(c("WEI(3,30)","WEI(1,50)","WEI(2.5,15)","WEI(5,20)","WEI(0.25,50)"),
        each=length(ts)))

ggplot(df,aes(x=t, y=y, color=fun)) +
    geom_line() +
    labs(y="density") +
    ylim(0,10)
```

The Weibull series system's hazard function is given by
\begin{equation}
\label{eq:sys_weibull_failure_rate_function}
h(t;\v\theta) =
    \sum_{j=1}^{m} \frac{k_j}{\lambda_j}\biggl(\frac{t}{\lambda_j}\biggr)^{k_j-1},
\end{equation}
whose proof follows from Theorem \ref{thm:sys_failure_rate}.

In Figure \@ref(series_haz), we plot the hazard function of the Weibull series
system with the component lifetime parameters considered earlier,
\begin{align*}
T_{i 1} &\sim \operatorname{WEI}(3,30)\\
T_{i 2} &\sim \operatorname{WEI}(2,50)\\
T_{i 3} &\sim \operatorname{WEI}(0.5,15)\\
T_{i 4} &\sim \operatorname{WEI}(5,20)\\
T_{i 5} &\sim \operatorname{WEI}(0.25,50)\\
\end{align*}
for $i=1,\ldots,n$.

```{r series_haz, fig.cap="Weibull series system hazard function plot.",fig.width=4,fig.asp=0.8,echo=F,warning=F,message=F}
h.series <- Vectorize(function(t) sum(h.wei1(t),h.wei2(t),h.wei3(t),
    h.wei4(t),h.wei5(t)))
df.series.haz <- data.frame(
    t=ts,
    y=h.series(ts))
ggplot(df.series.haz,aes(x=t, y=y)) +
    geom_line() +
    labs(y="hazard") +
    ylim(0,10)
```

The hazard function of a Weibull series system with the parameters for the
Weibull component lifetimes considered earlier is initially decreasing, then at
around $t=0.25$ it starts increasing (almost linearly).
This is a pattern we see in nature:

- High "infant mortality" rate.

- Wears out over time, an aging process.

The pdf of the series system is given by
\begin{equation}
\label{eq:sys_weibull_pdf}
f(t;\v\theta) =
\biggl\{
    \sum_{j=1}^m \frac{k_j}{\lambda_j}\left(\frac{t}{\lambda_j}\right)^{k_j-1}
\biggr\}
\exp
\biggl\{
    -\sum_{j=1}^m \bigl(\frac{t}{\lambda_j}\bigr)^{k_j}
\biggr\}.
\end{equation}
\begin{proof}
By definition,
$$
f(t;\v\theta) = h(t;\v\theta) R(t;\v\theta).
$$
Plugging in the failure rate and reliability functions given respectively by
Equations \eqref{eq:sys_weibull_reliability_function} and
\eqref{eq:sys_weibull_failure_rate_function} completes the proof.
\end{proof}

The conditional pmf of $K_i$ given $\T_i$ is given by
\begin{equation}
\label{eq:weibull_pmf_k_given_s}
f(k;\v\theta|t) = \frac{\frac{k_k}{\lambda_k}\bigl(\frac{t}{\lambda_k}\bigr)^{k_k-1}}
    {\sum_{j=1}^m \frac{k_j}{\lambda_j}\bigl(\frac{t}{\lambda_j}\bigr)^{k_j-1}}.
\end{equation}
\begin{proof}
By Theorem \ref{thm:f_given_s_form_2},
$$
f(k;\v\theta|t) = \frac{h_k(t;\v{\theta_k})}{\sum_{j=1}^m h_j(t;\v{\theta_j})}
$$
where $h_1,\ldots,h_m$ are the failure rate functions of the $m$ Weibull
component lifetimes.
\end{proof}

The joint pdf of $K_i$ and $\T_i$ is given by
\begin{equation}
\label{eq:weibull_joint_k_s}
f(k,t;\v\theta) = \frac{k_k}{\lambda_j}\biggl(\frac{t}{\lambda_k}\biggr)^{k_k-1}
\exp\biggl\{-\sum_{j=1}^{m}\biggl(\frac{t}{\lambda_j}\biggr)^{k_j}\biggr\}.
\end{equation}
\begin{proof}
Plugging in the conditional probability and the marginal probability given
respectively by Equations \eqref{eq:expo_prob_K_given_S} and
\eqref{eq:expo_sys_pdf} completes the proof.

By Theorem \ref{thm:f_k_and_t}, the joint pdf of $K_i$ and $\T_i$ is given by
$$
f(k,t;\v\theta) = h_k(t;\v{\theta_k}) R_{\T_i}(t;\v\theta)
$$
where $R_{T_i}(t;\v\theta)$ is given by Equation
\eqref{eq:sys_weibull_reliability_function} and $h_k$ is the failure rate
function of the $k$\textsuperscript{th} Weibull component.
\end{proof}

## Maximum likelihood estimation
The log-likelihood for our masked data model when we assume Conditions
\ref{cond:c_contains_k}, \ref{cond:equal_prob_failure_cause}, and
\ref{cond:masked_indept_theta} is given by
\begin{equation}
\ell(\v\theta) = \sum_{i=1}^n
\biggl\{
    \sum_{j=1}^m
    \bigl(\frac{s_i}{\lambda_j}\bigr)^{k_j}
\biggr\} +
\sum_{i=1}^n (1-\delta_i) \log
\Biggl\{
    \sum_{j \in c_i} \frac{k_j}{\lambda_j}\biggl(\frac{s_i}{\lambda_j}\biggr)^{k_j-1}
\Biggr\}.
\end{equation}
\begin{proof}
By Equation \eqref{eq:loglike}, 
$$
\ell(\v\theta) = \sum_{i=1}^n \log R(s_i;\v\theta) + \sum_{i=1}^n (1-\delta_i)
    \log \biggl\{ \sum_{k\in c_i} h_k(s_i;\v{\theta_k}) \biggr\}.
$$
Plugging in the component hazard system reliability functions given
respectively by Equations \eqref{eq:weibull_haz} and
\eqref{eq:sys_weibull_reliability_function} and simplifying completes the proof.
\end{proof}

We may find an MLE by solving the maximum likelihood equation \eqref{eq:mle_eq},
i.e., a  point $\v{\hat\theta} = (k_1,\lambda_1,\ldots,k_m,\lambda_m)$ satisfying
$$
\frac{\partial \ell}{\partial k_j}\Biggr|_{\hat{k}_j} = 0
$$
and
$$
\frac{\partial \ell}{\partial \lambda_j}\Biggr|_{\hat\lambda_j} = 0
$$
for $j=1,\ldots,m$ where
$\v{\hat\theta} = (\hat{k}_1,\hat\lambda_1,\ldots,\hat{k}_m,\hat\lambda_m)$.



$$
\frac{\partial \ell}{\partial \lambda_l} =
    \frac{k_l}{\lambda_l} \sum_{i=1}^n \biggl(\frac{s_i}{\lambda_l}\biggr)^{k_l} -
    \biggl(\frac{k_l}{\lambda_l}\biggr)^2 \sum_{i=1}^n 1_{\{l \in c_i\}}
    \biggl(\frac{s_i}{\lambda_l}\biggr)^{k_l-1}
    \Biggl\{\sum_{j \in c_i} \frac{k_j}{\lambda_j}\biggl(\frac{s_i}{\lambda_j}\biggr)^{k_j-1}\Biggr\}.
$$




$$
\frac{\partial \ell}{\partial k_l} =
    \frac{k_l}{\lambda_l} \sum_{i=1}^n \biggl(\frac{s_i}{\lambda_l}\biggr)^{k_l} -
    \biggl(\frac{k_l}{\lambda_l}\biggr)^2 \sum_{i=1}^n 1_{\{l \in c_i\}}
    \biggl(\frac{s_i}{\lambda_l}\biggr)^{k_l-1}
    \Biggl\{\sum_{j \in c_i} \frac{k_j}{\lambda_j}\biggl(\frac{s_i}{\lambda_j}\biggr)^{k_j-1}\Biggr\}.
$$



However, there is no closed-form solution for these equations, and so we
use the Newton-Raphson method described in Section \ref{sec:iterative},
where we also numerically approximate the gradient and Hessian using a finite
difference method.

## Real-world data set
As a trial run, we find the MLE on the masked data set provided by [@Huairu-2013].
See Appendix C to view the data.

This data set does not have right-censoring, but it does mask the component
cause of failure.
We do not know how any of this data was generated, but it is believed that
the component lifetimes can be adequately modeled by the Weibull distribution
and the candidate sets reasonable satisfy conditions $C_1$, $C_2$, and $C_3$.

We find the MLE and show its $95\%$-confidence interval with the following R code:
```{r eval=F}
guo.loglike <- md_loglike_weibull_series_C1_C2_C3(guo_weibull_series_md)
guo.theta.hat <- mle_newton_raphson(l=guo.loglike,theta0=rep(1,6))
point <- point(guo.theta.hat)
colnames(point) <- c("MLE")
cbind(point,confint(guo.theta.hat))
```

In Figure 5, we show a plot of the estimated parametric distribution of the series system
overlaid with a kernel density estimate of the observed lifetimes.

```{r density-plots, cache=T, fig.cap = "Estimated parametric pdf versus the kernel density plot of the observed lifetimes",echo=F,eval=F}
scales <- point(guo.theta.hat)[c(T,F,T,F,T,F)]
shapes <- point(guo.theta.hat)[c(F,T,F,T,F,T)]
t.pdf = function(t) dweibull_series(t,scales=scales,shapes=shapes)
lifetimes <- data.frame(t=guo_weibull_series_md$t)
ggplot(data.frame(t=lifetimes),aes(x=t)) +
    geom_density(aes(color="Kernel density estimate")) +
    geom_function(fun=t.pdf,aes(color="Theoretical density estimate")) +
    labs(x="t", y="density", color="Legend") +
    scale_color_manual(name="Type",
                       breaks=c('Kernel density estimate', 'Theoretical density estimate'),
                       values=c('Kernel density estimate'='blue', 'Theoretical density estimate'='green'))
```

These seem like a reasonable match. However, we can perform a hypothesis test
with
$$
    H_0 : \text{observed data is sampled from the estimated Weibull series system}.
$$
As we explained in Section \ref{sec:fit}, we use the Cramér-von Mises
goodness-of-fit test.

```{r,echo=F}
#library(goftest)
#gof <- cvm.test(x=guo_weibull_series_md$t,null=function(t) pweibull_series(t,scales=scales,shapes=shapes),estimated=T)
#p.value <- gof$p.value
p.value <- 1
```

Performing the test, we obtain a $p$-value of $`r p.value`$.
At the $95\%$ significance level, we conclude that the observed lifetimes
are compatible with the estimated Weibull series system.

## Simulation

Now, we wish to simulate a masked data set that meets Conditions
\ref{cond:c_contains_k}, \ref{cond:equal_prob_failure_cause}, and
\ref{cond:masked_indept_theta} by using the Bernoulli candidate model
with an appropriate choice of parameters.

We setup the parameters of this simulation with the following R code:
```{r}
# sample size
n <- 500

# m components in series system
m <- 3

# true parameter value of series system
theta <- c(2,40,      # lambda_1, k_1
           3,10,      # lambda_2, k_2
           2,30)      # lambda_3, k_3

# bernoulli candidate model parameter

# right censoring times
tau <- rep(2.05,n)
```

We generate a masked data sample of size $n=`r n`$, $m=`r m`$, $\tau_i=`r tau[1]`$,
$\v\theta=(`r theta`)'$, and using the simple Bernoulli candidate model 
consistent with conditions $C_1$, $C_2$, and $C_3$ with:
```{r,message=F,warning=F,eval=F}
md <- tibble(t1=stats::rweibull(n,shape=theta[2],scale=theta[1]),
             t2=stats::rweibull(n,shape=theta[4],scale=theta[3]),
             t3=stats::rweibull(n,shape=theta[6],scale=theta[5])) %>%
    md_series_lifetime() %>%
    md_series_lifetime_right_censoring(tau) %>%
    md_bernoulli_candidate_C1_C2_C3(m)
```

See Appendix A for more information about the `md_*` functions in the above
code.
The above code produces a data frame `md` whose first $6$ rows are given by
Table 2.

```{r echo=F,eval=F}
# "Simulated masked data for Weibull component lifetimes"
md.tmp <- md
md.tmp$x1 <- as.integer(md.tmp$x1)
md.tmp$x2 <- as.integer(md.tmp$x2)
md.tmp$x3 <- as.integer(md.tmp$x3)
md.tmp$delta <- as.integer(md.tmp$delta)
head(round(md.tmp,digits=3))
```

Now, we find the MLE with the following R code:
```{r eval=F}
loglike.weibull <- md_loglike_weibull_series_C1_C2_C3(md)
theta.hat <- mle_newton_raphson(
    l=loglike.weibull,
    theta0=theta)

points <- cbind(point(theta.hat),as.matrix(theta))
colnames(points) <- c("MLE","Parameter")
cbind(points,confint(theta.hat))
```

We see that they are reasonable estimates.

See Appendix B for more information about the `mle_newton_raphson` function in
the above code.


### Candidate set that generates a non-unique MLE

```{r eval=F}
md_block_candidate_m3 <- function(md)
{
    block <- function(k)
    {
        if (k == 1)
            return(c(T,T,F))
        if (k == 2)
            return(c(T,T,F))
        if (k == 3)
        {
            if (runif(1) < 0.1)
                return(c(T,T,T))
            else
                return(c(F,F,T))
        }
    }

    n <- nrow(md)
    x <- matrix(nrow=n,ncol=3)
    for (i in 1:n)
        x[i,] <- block(md$k[i])

    x <- tibble::as_tibble(x)
    colnames(x) <- paste0("x",1:3)
    md %>% dplyr::bind_cols(x)
}
```

```{r eval=F}
md.nu <- tibble(t1=stats::rweibull(n,shape=theta[2],scale=theta[1]),
                t2=stats::rweibull(n,shape=theta[4],scale=theta[3]),
                t3=stats::rweibull(n,shape=theta[6],scale=theta[5])) %>%
    md_series_lifetime() %>%
    md_series_lifetime_right_censoring(tau) %>%
    md_block_candidate_m3()
```


```{r eval=F,echo=F}
# Simulated masked data for Weibull component lifetimes
md.nu.tmp <- md.nu
md.nu.tmp$x1 <- as.integer(md.nu.tmp$x1)
md.nu.tmp$x2 <- as.integer(md.nu.tmp$x2)
md.nu.tmp$x3 <- as.integer(md.nu.tmp$x3)
md.nu.tmp$delta <- as.integer(md.nu.tmp$delta)
head(round(md.nu.tmp,digits=3),n=15)
```

Now, we find the MLE with the following R code:
```{r eval=F}
loglike.nu.weibull <- md_loglike_weibull_series_C1_C2_C3(md.nu)
theta.nu.hat <- mle_newton_raphson(
    l=loglike.nu.weibull,
    theta0=theta)

points.nu <- cbind(point(theta.nu.hat),as.matrix(theta))
colnames(points.nu) <- c("MLE","Parameter")
cbind(points.nu,confint(theta.nu.hat))
```

